---
title: \vspace{-1.5cm} **IDS702 Team Yellow Project I**
author: Anna Dai, Athena Liu, Dauren Bizhanov, Himangshu Ray Bhantana, Moritz Wilksch
date: October 07, 2021
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(ggplot2)
library(MASS)
library(rms)
library(readr)
library(GGally)
library(knitr)
library(xtable)
options(xtable.comment = FALSE, xtable.floating = FALSE)
library(kableExtra)
library(arm)
library(pROC)
library(e1071)
library(caret)
library(gridExtra)
library(dplyr)
library(stargazer)
library(car)

```


```{r, echo = FALSE, include=FALSE}
#import data files and clean data
rm(list = ls())

data <- read.csv("../Data/lalondedata.txt", sep=',')

data$treat <- as.factor(data$treat)
data$black <- as.factor(data$black)
data$hispan <- as.factor(data$hispan)
data$married <- as.factor(data$married)
data$nodegree <- as.factor(data$nodegree)
data$diff_re <- data$re78 - data$re75
```

# Summary

In this study, we examine whether there is evidence that employees who receive job training earn better wages than those who did not receive training. We use linear regression to study the relationship between earning differences and variables such as job training, educational background, and racial identity. We find that while the job training itself does not have a significant effect on the earnings difference between 1974 and 1978 at zero-age, it significantly affects the relationship that age has on the earnings difference. For treated individuals, an increase in age is associated with an increase in earnings difference, while the opposite is true for untreated individuals. This trend is not affected by race, years of schooling, a lack of a degree, or marital status. Overall, it is reasonable to conclude that the treatment's impact differs mostly by age group and not by any other demographic variables.  
To not just study the pre and post-training difference in income, we use a logistic regression model to examine whether job training is associated with a higher odds ratio of having a non-zero income in 1978. We find that treatment at age zero, being black, and an increase in age (without treatment) are all associated with lower odds of earning a non-zero income, whereas the pre-training income and an increase in age for treated people are associated with higher odds.

# Introduction

# Data
